---
title: "Sheet 5"
author: "Carsten Stahl"
date: "2023-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("drc")
```

### Exercise 16
We first import the data
```{r}
load("./data/data_gradedsheet.Rdata")

head(data_gradedsheet)
```
In this experiment we observe the effect of a herbicide on the length of leaves and the number of dead flies.

We see the variable "lenghth" is continuous and the "num.Dead" is binomial with the total of flies given in "total".

```{r}
leaves.4pLL <- drm(Length ~ Conc,
                   data = data_gradedsheet,
                   fct = LL2.4())

summary(leaves.4pLL)
```

We can see that the slope is decreasing (b > 0) with the upper asymptote at d  = 19,09 and the lower asymptote around c = 7,877.

Also the b parameter is only slightly different than zero, giving some evidence, that the effect is not as strong as the model tries to replicate.

e is the logistic ED50 value which lies around 1,3.

plotting the function leads to:
```{r}
plot(leaves.4pLL)
```

Here we can see the flat curve that the b parameter is indicating.

### ploting residuals
```{r}
plot(leaves.4pLL$predres)
```

We can see that the residuals slightly vary in their variance. This gives us reason to apply some method of dealing with variance heterogeneity.

```{r}
library(sandwich)
library(lmtest)
# fixing naming conflicts
estfun.drc <- drc::estfun.drc
bread.drc <- drc::bread.drc

# calculation of sandich estimation
coeftest(leaves.4pLL, vcov. = sandwich)
```
The standard-error for b has fallen, which points to a previously biased estimate. The p-value thus improves.

estimating after box-cox yields:
```{r}
leaves.4pLL.box.cox <- boxcox(leaves.4pLL, method ="anova", plot = F)

summary(leaves.4pLL.box.cox)
``` 
sigificance levels do not change and the estimates for the parameter do change, because we transformed the data.

### calculation of ED values
```{r}
ED(leaves.4pLL, c(10, 20, 80), type="relative")
```

Because of log-scaling we get negative values. these lie somewhere lower than 1. The relative ed values represent dosages where a specific percentage (10,20,80) of the maximal effect is surpassed.

### calculations of the bmd values
using the bmd function

```{r}


```
