---
title: "Sheet 3 Tox"
output: html_document
date: "2023-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 7
```{r}
library("drc")
```
first we need to load the data-set "terbuthylazin".
```{r}
data(terbuthylazin)
terbuthylazin
```
The dataset seems to have 8*3 dosages + 6 Controll dosages. The variable 'rgr' represents the growth rate on a type of lentil.

next let's fit the weibull model:
```{r}
terb.model.weibull.1 <- drm(rgr~dose,
                            data = terbuthylazin,
                            type = "continuous",
                            fct = weibull1())
terb.model.weibull.1
```
as help let's use the plot function to display this model:
```{r}
plot(terb.model.weibull.1)
```
First off the slope d is positive, therefore we have a descending curve.
Therefore we have d as the upper asymptote and c as the lower asymptote.
Furthermore the inflection point is more left  and lower. 


Let's fit with the lower asymptote fixed at 0:
```{r}
terb.model.weibull.1.fixed <- drm(rgr~dose,
                                  data = terbuthylazin,
                                  type = "continuous",
                                  fct = weibull1(
                                    fixed = c(NA, 0,NA, NA)
                                  ))
terb.model.weibull.1.fixed
```
Now the upper asymptote is a little bit higher while the slope decreased.
```{r}
plot(terb.model.weibull.1.fixed)
```

We can see here that the response for the highest dosage does not quite fit.
```{r}
ED(terb.model.weibull.1,
   c(10,20,50,80),
   type="relative")

ED(terb.model.weibull.1.fixed,
   c(10,20,50,80),
   type="relative")
```
Here we can see that the ED values for higher effects are much lower for the fixed model.

```{r}
e.max.weibull1 <- terb.model.weibull.1$coefficients[3] -terb.model.weibull.1$coefficients[2]
e.max.weibull1.fixed <- terb.model.weibull.1.fixed$coefficients[2]


print("for the unrestricted model:")
# print corresponding effects for the unrestricted model
for(i in c(10,20,50,80)) {
  print(terb.model.weibull.1$coefficients[3] - e.max.weibull1*(i/100))
}

print("for the restricted model:")
# print corresponding effects for the unrestricted model
for(i in c(10,20,50,80)) {
  print(e.max.weibull1*(1-i/100))
}
```

For the restricted model we see also lower effects for the ED-values.

Lets now plot the models together
```{r}
plot(terb.model.weibull.1)
plot(terb.model.weibull.1.fixed, add= T)
```
## Exercise 8
Let's first load the dataset
```{r}
data(algae)
algae
```
So 2 samples for 6 dosages and one control-group
Let's fit a 4pLL model:
```{r}
algae.4pLL <- drm(vol ~ conc,
                  data = algae,
                  fct = LL.4(),
                  type = "continuous")

algae.4pLL
```
Since b is positive we see a negative slope. We also know, that the lower asymptote is at -1.4 (which is not possible). Also the upper asymptote is at an effect size of 105 volume units.

```{r}
plot(algae.4pLL)
points(algae$conc,algae$vol)
```

Here we can see the slope of the model

```{r}
plot(algae$conc, residuals(algae.4pLL))
```
We can see, that the variance significantly decreases the higher the dosage. We therefore expect biased estimates for the variance of the residuals. 

Let's use the sandwich estimator to fix this bias.
```{r}
# avoiding naming conflicts
estfun.drc <- drc::estfun.drc
bread.drc <- drc::bread.drc

sandwich::sandwich(algae.4pLL)
```