---
title: "Sheet 3 Tox"
output: html_document
date: "2023-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 7
```{r}
library("drc")
```
first we need to load the data-set "terbuthylazin".
```{r}
data(terbuthylazin)
terbuthylazin
```
The dataset seems to have 8*3 dosages + 6 Controll dosages. The variable 'rgr' represents the growth rate on a type of lentil.

next let's fit the weibull model:
```{r}
terb.model.weibull.1 <- drm(rgr~dose,
                            data = terbuthylazin,
                            type = "continuous",
                            fct = weibull1())
terb.model.weibull.1
```
as help let's use the plot function to display this model:
```{r}
plot(terb.model.weibull.1)
```
First off the slope d is positive, therefore we have a descending curve.
Therefore we have d as the upper asymptote and c as the lower asymptote.
Furthermore the inflection point is more left  and lower. 


Let's fit with the lower asymptote fixed at 0:
```{r}
terb.model.weibull.1.fixed <- drm(rgr~dose,
                                  data = terbuthylazin,
                                  type = "continuous",
                                  fct = weibull1(
                                    fixed = c(NA, 0,NA, NA)
                                  ))
terb.model.weibull.1.fixed
```
Now the upper asymptote is a little bit higher while the slope decreased.
```{r}
plot(terb.model.weibull.1.fixed)
```

We can see here that the response for the highest dosage does not quite fit.
```{r}
ED(terb.model.weibull.1,
   c(10,20,50,80),
   type="relative")

ED(terb.model.weibull.1.fixed,
   c(10,20,50,80),
   type="relative")
```
Here we can see that the ED values for higher effects are much lower for the fixed model.

```{r}
e.max.weibull1 <- terb.model.weibull.1$coefficients[3] -terb.model.weibull.1$coefficients[2]
e.max.weibull1.fixed <- terb.model.weibull.1.fixed$coefficients[2]


print("for the unrestricted model:")
# print corresponding effects for the unrestricted model
for(i in c(10,20,50,80)) {
  print(terb.model.weibull.1$coefficients[3] - e.max.weibull1*(i/100))
}

print("for the restricted model:")
# print corresponding effects for the unrestricted model
for(i in c(10,20,50,80)) {
  print(e.max.weibull1*(1-i/100))
}
```

For the restricted model we see also lower effects for the ED-values.

Lets now plot the models together
```{r}
plot(terb.model.weibull.1)
plot(terb.model.weibull.1.fixed, add= T)
```
## Exercise 8
Let's first load the dataset
```{r}
data(algae)
algae
```
So 2 samples for 6 dosages and one controll-group
Let's fit a 4pLL model:
