---
title: "sheet 2 solutions (coding part)"
author: "carsten stahl"
date: "2023-10-17"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the data
Before starting this exercise we are going to load the data and requirements

```{r}
library(tidyverse)
library(nlme)
load("./data/SimulatedTreatmentEffect.RData")
```

## a)

```{r}
gnls_treatment <- gnls(resp ~ th1+(th4-th1)/(1+(exp((conc-th2)*th3))),
                       data = conc.resp.df,
                       params=list(th1+th2+th3+th4~1),
                       control=gnlsControl(nlsTol=0.1),
                       start=c(0, 2, 1, 100))

summary(gnls_treatment)
```
- The slope is positive -> upward slope
- The response ranges from 2 to almost 100
- The alert-dosage (logscaled) is at around 2 (a little smaller than 1)

## b)

```{r}
# creating dummy variables
conc.resp.df$in_T1 <- conc.resp.df$treat == "T1"
conc.resp.df$in_T2 <- conc.resp.df$treat == "T2"

# applying model
gnls_treatment_with_dummy <- gnls(resp~(th1+(th4-th1)/(1+(exp((conc-th2)*th3))))*in_T1 +
                                    (th1+(th4-th1)/(1+(exp((conc-th2)*th3))))*in_T2,
                       data = conc.resp.df,
                       params=list(th1+th2+th3+th4~1),
                       control=gnlsControl(nlsTol=0.1),
                       start=c(0, 2, 1, 100))

summary(gnls_treatment_with_dummy)
```

Nothing really changed

## c)
```{r}
gnls_treatment_with_dummy_diff <- gnls(resp~(th1+(th4-th1)/(1+(exp((conc-th21)*th31))))*in_T1 +
                                    (th1+(th4-th1)/(1+(exp((conc-th22)*th32))))*in_T2,
                                 data = conc.resp.df,
                                 params=list(th1+th21+th22+th31+th32+th4~1),
                                 control=gnlsControl(nlsTol=0.1),
                                 start=c(0, 2, 2, 1, 1, 100))

summary(gnls_treatment_with_dummy_diff)

```

Both alert-concentration and slope have changed after treatment. The model is expected to have a better fit.

## d)
```{r}
anova(gnls_treatment_with_dummy_diff, gnls_treatment)
```

Yes, the p value is suffiently low to say, that the models are different.
