---
title: "Sheet 4"
output: html_document
date: "2023-12-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("drc")
```

### Exercise 11
This exercise is all about binomial response-data.

```{r}
load("./data/Binomial-DoseResponse.Rdata")

binom.3pLL <- drm(events ~ dose,
                  fct = LL.3u(),
                  data = binom.events,
                  type = "binomial",
                  weights = total,
                  curveid = treat)
binom.3pLL
```
```{r}
plot(binom.events$dose, binom.events$events)
```
Plotting the models

```{r}
plot(binom.3pLL)
```

```{r}
ED(binom.3pLL, 20, type="relative")
```

### Exercise 12
Now counting data
```{r}
load("./data/Nrleaves.Rdata")

leaves.model <- drm(NrLeaves ~ Concentration,
                    data = Nrleaves.df,
                    fct = LL.4(),
                    type = "Poisson")

summary(leaves.model)
```

We can see: 
- b is positive -> negative slope -> compounds inhibits leaf growth
- c the lower asymptote is at 2.75 leaves
- d the upper asymptote is at 18.9 leaves
- e ED50 dosage at 1.12

let's plot the model just to be sure
```{r}
plot(leaves.model)
```

Here we can see the decreasing slope.

### including times
we are going to include time as a weight
```{r}
leaves.model.with.weigths <- drm(NrLeaves ~ Concentration,
                                data = Nrleaves.df,
                                fct = LL.4(),
                                weights = Days,
                                type = "Poisson")


summary(leaves.model.with.weigths)
```
Estimation is now normalized to day. Therefore the fitted values are no longer in sync with the data.
```{r}

```

### Exercise 13
### b)
We only have given that $X\sim \mathcal N(\mu,\sigma^2)$.

We can apply taylor approximation. 
### c)
same with variance 

### Exercise 14
loading the bmd package
```{r}
# im just going to write down the code and not execute it
# (i)
bmd(binom.3pLL, bmr=0.1, backgType = "modelbased", def="additional")
# (ii)
bmd(binom.3pLL, bmr = 0.1, backgType = "modelbased", def = "excess")
# (iii)
bmd(binom.3pLL, bmr = 0.1, backgType = "absolute", backg = 0.15, def = "additional")
# (iv)
bmd(binom.3pLL, bmr = 0.1, backgType = "absolute", backg = 0.15, def = "excess")

## b)
#(i)
bmd(leaves.4pLL, bmr = 0.1, backgType = "absolute", backg = 85, def = "hybridAdd")
#(ii)
bmd(leaves.4pLL, bmr = 0.1, backgType = "absolute", backg = 85, def = "hybridExc")

#(iii)
bmd(leaves.4pLL, bmr = 0.1, backgType = "hybridSD", def = "hybridAdd")
# (iv)
bmd(leaves.4pLL, bmr = 0.1, backgType = "hybridSD", def = "hybridExc")
```