---
title: "sheet_1_ex_1"
output:
  html_document: default
  pdf_document: default
date: "2023-10-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
In the first lecture of this course we got introduced to a study about the **adverse effects of valporic acid on neurological functionality in embryos during pregnancy** (Kurg. A.K. et. a, (2013)). Valporic acid is typically used for treatment of epilepsy, migraines, headaches and even specific types of mental illnesses as a mood stabilizer. However, according to the World Health Organization, it is strongly advised to not take this medication during pregnancy as it bears a high risk of developmental disorders (more information, like if you have literally nothing else to do [here](https://www.who.int/news/item/02-05-2023-use-of-valproic-acid-in-women-and-girls-of-childbearing-potential#:~:text=Valproic%20acid%20%28sodium%20valproate%29%20should%20not%20be%20prescribed,to%20valproic%20acid%20%28sodium%20valproate%29%20in%20the%20womb.)). The study has been done in vitro (in petri dishes) on (multiplied) embryonic stem cells. As the **response-data** we therefore use **gen-expressions, ranging from 2 to 14**. Beside the **6 negative controls** we have seen **7 response-dosages on 3 samples each**.

The data-set has been reduced quite a bit: out of the 10.000 gen expressions, only 500 were chosen based on their variance across **all** samples.

## Requirements
As part of this analysis, we are using tidyverse (especially ggplot).
```{r}
library(tidyverse)
library(gridExtra)
```

## Importing the data
To import the data we use the `load()` function. Make sure to have the right working directory, when importing the data. 

```{r}
# use the setwd(...) function if necessary.
load("./data/VPAData-Random.Rda")
head(randomVPA)
```
We will refactor the data to have an easier time with the given tasks:
```{r}

# we will create the variables for gen expression, sample_id, concentration and gen id
# length of the data
n <- 27*500
data <- data.frame(gen_exp = rep(NA, n),
                   gen_id = rep(1:500, 27),
                   concentration = rep(NA, n),
                   sample_id = rep(NA, n))

# this part of the code is very slow and cheap. Feel free to find a more elegant solution
response_dose <- c(25, 150, 350, 450, 550, 800, 1000)

data$concentration[1:(6*500)] <- 0

for (i in 1:7) {
  lower_bound <- 6*500 + (i-1)*1500
  upper_bound <- 6*500 + i*1500
  data$concentration[lower_bound:upper_bound] <- response_dose[i]
}

sample_ids <- factor(as.character(1:27),
                     levels = as.character(1:27))

for (i in 1:27) {
  data$sample_id[((i-1)*500):(i*500)] <- sample_ids[i]
  for (j in 1:500) {
    coordinate_to_row <- (i-1)*500 + j
    data$gen_exp[coordinate_to_row] <- randomVPA[j,i]
    
  }
}
```
## Boxplot from the lecture
The plot from the lecture will be realized using ggplot2
```{r, out.width="90%", fig.align="center"}
data %>%
  ggplot(aes(y = gen_exp, x = sample_id, group = sample_id)) +
  geom_boxplot()

```

## Variance of genes after response-dosage
We are going to use tidyverse to calculate the variances. Afterwards using ggplot2 to plot the histogram.
```{r, out.width="90%", fig.align="center"}
# using new data set
var_data <- data %>%
  group_by(gen_id, concentration) %>%
  summarise(var_gen_exp = sd(gen_exp)) %>%
  unique()

# histograms by concentration
for (i in 1:8) {
  plot_name <- paste("hist_plot_", i,
                     sep = "")
  # filtering data and creating a fitting x label
  if (i == 1) {
    plot_data <- var_data %>%
      filter(concentration == 0)
    plot_x_label <- paste("response-dosage:", 0)
  } else {
    plot_data <- var_data %>%
      filter(concentration == response_dose[i-1])
    plot_x_label <- paste("response-dosage:", response_dose[i-1])
  }
  
  # creating a plot and assigning it a name
  assign(plot_name,
         plot_data %>%
           ggplot(aes(x = var_gen_exp)) +
           geom_histogram() +
           xlim(0,1) +
           xlab(plot_x_label))
}
grid.arrange(hist_plot_1,
             hist_plot_2,
             hist_plot_3,
             hist_plot_4,
             hist_plot_5,
             hist_plot_6,
             hist_plot_7,
             hist_plot_8,
             ncol = 2
             )
```

## Ploting the gene profiles with the highest variance
For this we are going to use `var_data` again
```{r, out.width="90%", fig.align="center"}
# finding the max values after concentration
max_var_data <- var_data %>%
  group_by(concentration) %>%
  summarise(max_var = max(var_gen_exp),
            # find the gen id that corresponds to the maximum value
            gen_id = gen_id[which(var_gen_exp == max(var_gen_exp))]) %>%
  unique()

# inner join with processed data
max_var_plot_data <- data %>%
  filter(gen_id %in% max_var_data$gen_id)

# plotting
max_var_plot_data %>% ggplot(aes(x = sample_id,
                                 y = gen_exp,
                                 group = gen_id)) +
  geom_line(size = 1)
```

afterwards we are going to repeat the same procedure for min values:
```{r, out.width="90%", fig.align="center"}
# finding the max values after concentration
min_var_data <- var_data %>%
  group_by(concentration) %>%
  summarise(min_var = min(var_gen_exp),
            # find the gen id that corresponds to the maximum value
            gen_id = gen_id[which(var_gen_exp == min(var_gen_exp))]) %>%
  unique()

# inner join with processed data
min_var_plot_data <- data %>%
  filter(gen_id %in% min_var_data$gen_id)

# plotting
min_var_plot_data %>% ggplot(aes(x = sample_id,
                                 y = gen_exp,
                                 group = gen_id)) +
  geom_line(size = 1)
```